On verifying AhirV2 generated VHDL using software testbenches

### 2012-03-04 v2/CtestBench/docs/CtestBench.tex
Added support for stacks in llvm2aa (through pipedepth files).  Also modified the Queue/SynchFifo to try to merge pipes and stacks in the same entity.. but it is getting messy and it may be best to split the two.

\end{document}

>>>>

\section{An example}

Consider the following program (lets say it is in file ``prog.c''):
\begin{verbatim}
#include <stdlib.h>
#include <stdint.h>
#include <stdio.h>
// this file provides definitions
// of pipe access functions read/write.
#include <Pipes.h>

// note: initialized value..
uint32_t sum1 = 23;
uint32_t sum2 = 39;

// note: no problems with pointers :-)
uint32_t* tgt[2] = {&sum1, &sum2};

void accumulate()
{
    int i = 0;
    while(1)
    {
	int nxt = read_uint32("in_data");
#ifdef SW
	printf("read %u\n", nxt);
#endif
	// ugly, but this is just a demo!
	*(tgt[i])= ((*tgt[i]) + nxt);

	write_uint32("out_data",*(tgt[i]));
#ifdef SW
	printf("wrote %u\n", *(tgt[i]));
#endif
	i = 1 - i;
    }
}

uint32_t get_sum(uint32_t idx)
{
    return(*(tgt[idx]));
}

\end{verbatim}

This program describes a {\em system} which listens
for data on a pipe ``in\_data'', and sends data
out on a pipe ``out\_data''.  The incoming data
is accumulated into the variable {\em sum}, and
there are two methods to set and get the value
of {\em sum}.

Now to test this program, we can write a test-bench
such as this one (lets call this file ``testbench.c'').
\begin{verbatim}
#include <pthread.h>
#include <signal.h>
#include <stdio.h>
#include <stdlib.h>
#include <stdint.h>

#ifdef SW
// for the pipe read/write functions
#include <Pipes.h>
// for the pipeHandler daemon..
#include <pipeHandler.h>
#include "prog.h"
#else
#include "vhdlCStubs.h"
#endif

void Exit(int sig)
{
	fprintf(stderr, "## Break! ##\n");
	exit(0);
}

void *accumulate_(void* fargs)
{
   accumulate();
}

void *write_pipe_(void* a)
{
   write_uint32_n("in_data",(uint32_t*)a, 10);
}

void *read_pipe_(void* a)
{
   read_uint32_n("out_data",(uint32_t*)a, 10);
}

void* pipeHandler__(void* a)
{
	pipeHandler();
}

int main(int argc, char* argv[])
{
	signal(SIGINT,  Exit);
  	signal(SIGTERM, Exit);

	uint32_t data_in[10], data_out[10];
        int i;

#ifdef SW

	// in the "software" case, we start a pipe-handler
	// to manage the pipes..  Earlier, we were using named
	// pipes, which were very flaky and difficult to debug.
	// the pipeHandler generates a log file (pipeHandler.log)
	// which is very useful for figuring out what happened.
	// one can also use gdb to trace activity.
	pthread_t phandler_t;
	pthread_create(&phandler_t,NULL,&pipeHandler__,NULL);

	usleep(100);

	// register two FIFOs..
	register_pipe("in_data",10,32,0);
	register_pipe("out_data",10,32,0);

#endif

#ifndef SW
	// to set the initial value of sum.
	// in the hardware version, storage
	// variables are initialized by calling
	// this function (auto-generated by
	// the Aa linker AaLinkExtMem)
	global_storage_initializer_();
#endif

        for(i = 0; i < 10; i++)
	{
		data_in[i] = i;
	}

	pthread_t acc_t, wpipe_t, rpipe_t;

#ifdef SW

	pthread_create(&acc_t,NULL,&accumulate_,NULL);
#endif
	pthread_create(&wpipe_t,NULL,&write_pipe_,(void*)data_in);
	pthread_create(&rpipe_t,NULL,&read_pipe_,(void*)data_out);

	pthread_join(wpipe_t,NULL);
	pthread_join(rpipe_t,NULL);

 	fprintf(stdout,"from out_data, we read ");
	for(i=0; i < 10; i++)
 	  fprintf(stdout," %u ", data_out[i]);
	fprintf(stdout,"\n");

	fprintf(stdout,"Sum 0 is %d\n",get_sum(0));
	fprintf(stdout,"Sum 1 is %d\n",get_sum(1));

#ifdef SW
	pthread_cancel(acc_t);
	killPipeHandler();
#endif
}
\end{verbatim}
The test-bench starts
three threads: one to write data to pipe in\_data, one
to read data from pipe out\_data, and one to run the
accumulate function.  After the last data is read back
from out\_data, the test-bench gets the value of sum
and prints it out.

Obviously, we would prefer to use the same test-bench
to verify that the VHDL system generated from ``prog.c'' 
functions correctly.  
The difference is that instead of using the pipeHandler,
the test-bench now uses methods in SocketLib.  Further, the
VHDL is executed in a VHDL simulator; the simulator communicates
with the testbench using sockets.
The {\em ifdef's} in the test-bench and the system program
indicate the difference between the pure software version
of the system-test-bench combination and the hardware-software
version.

The following Makefile builds a software-only testbench
executable, and also converts the system described in prog.c
to VHDL.  The same testbench can be used to test the VHDL
also.
\begin{verbatim}
SOCKETLIB_INCLUDE=../../CtestBench/include
SOCKETLIB_LIB=../../CtestBench/lib
VHDL_LIB=../../vhdl
VHDL_VHPI_LIB=../../CtestBench/vhdl
all: SW HW

# compile with SW defined... this
# tests the SW.
SW: prog.c prog.h testbench.c 
	gcc -c -DSW -I$(SOCKETLIB_INCLUDE) prog.c
	gcc -c -DSW -I$(SOCKETLIB_INCLUDE) testbench.c
	gcc -o testbench_sw prog.o testbench.o -L$(SOCKETLIB_LIB) -lSocketLib -lpthread

# five steps from C to vhdl simulator.
HW: c2llvmbc llvmbc2aa aa2vc vc2vhdl vhdlsim

# C to llvm byte-code.. use clang.
c2llvmbc: prog.c prog.h testbench.c
	clang -std=gnu89 -I$(SOCKETLIB_INCLUDE) -emit-llvm -c prog.c
	llvm-dis prog.o

# llvm byte-code to Aa..
llvmbc2aa:  prog.o
	llvm2aa --storageinit=true prog.o | vcFormat >  prog.aa

# Aa to vC
aa2vc: prog.aa
	AaLinkExtMem prog.aa | vcFormat > prog.linked.aa
	Aa2VC -O -C prog.linked.aa | vcFormat > prog.vc

# vC to VHDL
vc2vhdl: prog.vc
	vc2vhdl -C -e ahir_system -w -s ghdl -T accumulate\
               -t get_sum -t global_storage_initializer_ -f prog.vc 
	vhdlFormat < ahir_system.unformatted_vhdl > ahir_system.vhdl
	vhdlFormat < ahir_system_test_bench.unformatted_vhdl\
                       > ahir_system_test_bench.vhdl

# build testbench and ghdl executable
# note the use of SOCKETLIB in building the testbench.
vhdlsim: ahir_system.vhdl ahir_system_test_bench.vhdl testbench.c vhdlCStubs.h vhdlCStubs.c
	gcc -c vhdlCStubs.c -I./ -I$(SOCKETLIB_INCLUDE)
	gcc -c testbench.c -I./ -I$(SOCKETLIB_INCLUDE)
	gcc -o testbench_hw testbench.o vhdlCStubs.o  -L$(SOCKETLIB_LIB) -lSocketLib -lpthread
	ghdl --clean
	ghdl --remove
	ghdl -i --work=ahir  $(VHDL_LIB)/ahir.vhdl
	ghdl -i --work=ieee_proposed  $(VHDL_LIB)/ieee_proposed.vhdl
	ghdl -i --work=work $(VHDL_VHPI_LIB)/Utility_Package.vhdl
	ghdl -i --work=work $(VHDL_VHPI_LIB)/Vhpi_Package.vhdl
	ghdl -i --work=work ahir_system.vhdl
	ghdl -i --work=work ahir_system_test_bench.vhdl
	ghdl -m --work=work -Wc,-g -Wl,-L$(SOCKETLIB_LIB) -Wl,-lVhpi ahir_system_test_bench

clean:
	rm -f *.o* *.cf *.*vhdl vhdlCStubs.* *.vcd in_data* out_data* testbench_sw testbench_hw ahir_system_test_bench vhpi.log *.aa *.vc

PHONY: all clean	
\end{verbatim}
To test the software, run testbench\_sw.  To verify the hardware (using
the VHDL simulator GHDL), start testbench\_sw in one shell, and then start ahir\_system\_test\_bench
in a different shell.
\end{document}
